---
title: " RingdateR output log"
output:
  html_document:
    theme: cosmo
    toc: no
---
```{r run_info, echo = FALSE}

start_time <- Sys.time()

undated_data <- NULL
dated_chron <- NULL

# if (nrow(chron_detrended$df_data) < 1){
#  undated_data <- "No data loaded"
#  dated_chron <- "No data loaded"
# } else {
  undated_data <- input$file1$name
  
  dated_chron <- input$file2$name
# }

```

Run time: `r start_time`

Data loaded: `r undated_data`

Chronology loaded: `r dated_chron`

```{r det_mode, echo = FALSE}
  det_val <- as.numeric(input$detrending_select)  

  if (det_val == 1){
    det_method = "No detrending applied"
  } else if (det_val == 2){
    det_method <- "Convert to z-scores"
  } else if (det_val == 3){
    det_method <- paste0(input$splinewindow, " year spline")
  } else if (det_val == 4){
    det_method <- "Mod. negative exponential"
  } else if (det_val == 5){
    det_method <- "Friedman"
  } else if (det_val == 6){
    det_method <- "ModHugershoff"
  } else if (det_val == 7){
    det_method <- "First difference"
  } else {det_method <- "no number"}
```

### Detrending and report settings applied:
Detrending mode: `r det_method`

Prewhitening: `r input$ARmod`

Log Transform:`r input$logT`

Verbose :`r input$summary_verbose`

If verbose = TRUE: all possible leads and lags are evaluated.
If verbose = FALSE: lead-lag analysis is limited to a +/-20 year range of lags.

Problem sample window: `r input$summary_probs` years (with 50% overlap)


### Correlations between each series and the arithemtic mean chronology with replacement
(The sample being analysed is excluded from the chronology)

```{r table, echo = FALSE}
  the_data<-initiated_chron_correl_replace()
  row.names(the_data) <-1:nrow(the_data)
  knitr::kable(the_data, floating.environment="sidewaystable", digits = 32)
```

### Chronology and EPS plots

EPS and Rbar window: `r input$rbar_wind_init` years with 50% overlap

```{r chron_plot, echo = FALSE, out.width = '100%'}
  plot_initiated_chron(text.size = 8, 
                                 line.width = 0.5,
                                 plot.line = 0.5
                                 )
  
```
  

  
```{r problems, echo = FALSE}
  probs <- prob_check(final_chron_alinged$df_data,
                      as.numeric(input$summary_probs))
  row.names(probs) <- 1:nrow(probs)
  
```

### Distribution of aligned samples

```{r sample_dist, echo = FALSE, out.width = '100%', dpi = 120}
  Sample_dist_fun(ft_size = 8, 
                  line.width = 0.5, 
                  plot.line = 0.75, 
                  sample_id = 0)
                  
```

## Overview of correlations between each series and the arithmetic mean chronology
```{r problem_sample_plots, echo = FALSE, out.width = '100%', fig.out = 10, dpi = 120}
 
  full_data <- final_chron_alinged$df_data

 for (i in 2:ncol(full_data)){

  sample <- full_data[,i]
  ex_sample <- full_data[,-c(1,i)]
  years <- full_data[,1]
  chron<-rowMeans(ex_sample, na.rm = TRUE)
  combined<-data.frame(years, chron, sample)
  colnames(combined) <- c("years", "chronology", colnames(full_data)[i])

   test<-subset(combined, complete.cases(combined))
  
  if(nrow(test)>input$cor_win){
    hm_plot <- heatmap_analysis(combined, 
                                colnames(combined)[2], 
                                colnames(combined)[3], 
                                neg_lag = -5, 
                                pos_lag = 5, 
                                win = as.numeric(input$cor_win), 
                                center = 0, 
                                complete = FALSE, 
                                sel_col_pal = 1, 
                                font_size = 8, 
                                axis_line_width = 0.5, 
                                plot_line = 0.75, 
                                leg_size = 1)
  }
  
  chron_comp <- lead_lag_analysis(the_data = combined,
                                 mode = 1, neg_lag = -20, pos_lag = 20,
                                 complete = input$summary_verbose)

  combined<-subset(combined, complete.cases(combined))

  ln_plot <- line_plot(combined, colnames(combined)[2], colnames(combined)[3], lag = 0, text.size = 8, line.width = 0.5, plot.line = 0.5)
  
  bar_graph <- lead_lag_bar(the_data = as.data.frame(chron_comp[2]),
                  sample_1 = colnames(combined)[2],
                  sample_2 = colnames(combined)[3],
                  text.size = 8, 
                  line.width = 0.5)
  

  g1 <- ggplotGrob(ln_plot)
  g2 <- ggplotGrob(bar_graph)
  g3 <- ggplotGrob(hm_plot)
  

  g<-rbind(g1, g2, g3, size = "first")
  grid::grid.newpage()
  
  print(colnames(combined)[3])
  grid.draw(g)
  
  ######
  
    samp_names<-paste0("ser_1_",colnames(combined)[2], "_ser_2_",colnames(combined)[3], "_")
    master_names<-c("lag", "R_Val", "P_Val", "T_val", "Overlap", "First_ring", "Last_ring")
    master_names<-paste0(samp_names,master_names)

  if(!master_names[1] %in% colnames(as.data.frame(chron_comp[2]))){
    stop("Error in lead_lag_bar(): Resilts not found for selected samples")
  }
    matches <- c("Best match", "Second best", "Third best")
    lag <-as.data.frame(chron_comp[2])[[master_names[1]]]
    R_Val<-as.data.frame(chron_comp[2])[[master_names[2]]]
    P_Val <-as.data.frame(chron_comp[2])[[master_names[3]]]
    T_val <-as.data.frame(chron_comp[2])[[master_names[4]]]
    Overlap <-as.data.frame(chron_comp[2])[[master_names[5]]]
    First_ring	<-as.data.frame(chron_comp[2])[[master_names[6]]]
    Last_ring <-as.data.frame(chron_comp[2])[[master_names[7]]]
    selected<-data.frame(lag, R_Val, P_Val, T_val, Overlap, First_ring, Last_ring)

    # remove neg values
    selected<-subset(selected,(selected[,2]>0))

    ordered<-selected[order(selected[,3]),]
    ordered<-cbind(matches, ordered[1:3,])
    row.names(ordered)<-1:3
    
    print(knitr::kable(ordered, floating.environment="sidewaystable"))

   prob_found <- 0
    for (i in 1:nrow(probs)){
      if(probs[i,1] == colnames(combined)[3]){
        prob_found <- i
      } 
    }
    
    if(prob_found >=1){
        print(knitr::kable(probs[prob_found,],floating.environment="sidewaystable"))
      } else {
        print("Problem checker could not detect problems with this sample.")
      }
   
    
  combined<-NULL
 }

```

```{r run_time, echo = FALSE}

run_dur <- Sys.time() - start_time  

```

Run duration: `r run_dur` 
